---
title: "combine combustion data by project (FISL, LC, LCC"
author: "Anna Talucci"
date: "2025-06-06"
output: html_document
---

# Overview

Combine CSV files at project level for combustion model and for day of burn data.

# Packages 

```{r}
library(tidyverse)
library(sf)
```


# Data 

# Points

```{r}
lc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/lc_pts_WGS84_2025-06-02.shp", "lc_pts_WGS84_2025-06-02")
```

```{r}
lcc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/lcc_pts_WGS84_2025-06-02.shp", "lcc_pts_WGS84_2025-06-02")
```

```{r}
fisl_pts =st_read("../outputs/LCC_cleaned/points/shapefiles/fisl_pts_WGS84_2025-06-02.shp", "fisl_pts_WGS84_2025-06-02")
```

# CSV

## FISL

```{r}
fisl_indices_list = list.files("../data/combustion_data_layers/FISL/indices", pattern='.csv$', full.names = TRUE)
fisl_indices_list
```

```{r}
fisl_tct_list = list.files("../data/combustion_data_layers/FISL/tct", pattern='.csv$', full.names = TRUE)
fisl_tct_list
```

```{r}
fisl_treecover_list = list.files("../data/combustion_data_layers/FISL/treecover", pattern='.csv$', full.names = TRUE)
fisl_treecover_list
```

```{r}
fisl_static_list = list.files("../data/combustion_data_layers/FISL/static", pattern='.csv$', full.names = TRUE)
fisl_static_list
```

```{r}
fisl_dob_list = list.files("../data/combustion_data_layers/FISL/dob", pattern='.csv$', full.names = TRUE)
fisl_dob_list
```

```{r}
fisl_cna = read_csv("../data/combustion_data_layers/FISL/2025-06-02_climateNA_FISL_Normal_1981_2010MSY.csv")
```

## LC

```{r}
lc_indices_list = list.files("../data/combustion_data_layers/LC/indices", pattern='.csv$', full.names = TRUE)
lc_indices_list
```

```{r}
lc_tct_list = list.files("../data/combustion_data_layers/LC/tct", pattern='.csv$', full.names = TRUE)
lc_tct_list
```

```{r}
lc_treecover_list = list.files("../data/combustion_data_layers/LC/treecover", pattern='.csv$', full.names = TRUE)
lc_treecover_list
```

```{r}
lc_static_list = list.files("../data/combustion_data_layers/LC/static", pattern='.csv$', full.names = TRUE)
lc_static_list
```

```{r}
lc_dob_list = list.files("../data/combustion_data_layers/LC/dob", pattern='.csv$', full.names = TRUE)
lc_dob_list
```

```{r}
lc_cna = read_csv("../data/combustion_data_layers/LC/2025-06-02_climateNA_LC_Normal_1981_2010MSY.csv")
```

## LCC

```{r}
lcc_indices_list = list.files("../data/combustion_data_layers/LCC/indices", pattern='.csv$', full.names = TRUE)
lcc_indices_list
```

```{r}
lcc_tct_list = list.files("../data/combustion_data_layers/LCC/tct", pattern='.csv$', full.names = TRUE)
lcc_tct_list
```

```{r}
lcc_treecover_list = list.files("../data/combustion_data_layers/LCC/treecover", pattern='.csv$', full.names = TRUE)
lcc_treecover_list
```

```{r}
lcc_static_list = list.files("../data/combustion_data_layers/LCC/static", pattern='.csv$', full.names = TRUE)
lcc_static_list
```

```{r}
lcc_dob_list = list.files("../data/combustion_data_layers/LCC/dob", pattern='.csv$', full.names = TRUE)
lcc_dob_list
```

```{r}
lcc_cna = read_csv("../data/combustion_data_layers/LCC/2025-06-02_climateNA_LCC_Normal_1981_2010MSY.csv")
```

# Read list of files 

## Functions for csv list

```{r}
csv_rbind <- function(file_list) {
  csv_data <- purrr::map_dfr(file_list, readr::read_csv)
  return(csv_data)
}

```

```{r}
csv_cbind <- function(file_list, id_col = "ID") {
  
  # Read and join using reduce with left_join by ID
  data_list <- purrr::map(file_list, ~readr::read_csv(.x, show_col_types = FALSE))
  
  # Ensure ID column exists in all files
  data_list <- purrr::map(data_list, ~dplyr::select(., dplyr::all_of(id_col), dplyr::everything()))
  
  # Join all data frames by ID
  joined_data <- purrr::reduce(data_list, ~dplyr::full_join(.x, .y, by = id_col))
  
  return(joined_data)
}

```




## dob

```{r}
( fisl_dob = csv_rbind(fisl_dob_list) ) 
( lc_dob = csv_rbind(lc_dob_list) ) 
( lcc_dob = csv_rbind(lcc_dob_list) ) 
```

## indices

```{r}
( fisl_indices = csv_rbind(fisl_indices_list) ) 
( lc_indices = csv_rbind(lc_indices_list) ) 
( lcc_indices = csv_rbind(lcc_indices_list) ) 
```

## tree cover

```{r}
( fisl_treecover = csv_rbind(fisl_treecover_list) ) 
( lc_treecover = csv_rbind(lc_treecover_list) ) 
( lcc_treecover = csv_rbind(lcc_treecover_list) ) 
```


## tct

```{r}
( fisl_tct = csv_rbind(fisl_tct_list) ) 
( lc_tct = csv_rbind(lc_tct_list) ) 
( lcc_tct = csv_rbind(lcc_tct_list) ) 
```

# Static

```{r}
( fisl_static = csv_cbind(fisl_static_list))
( lc_static = csv_cbind(lc_static_list))
( lcc_static = csv_cbind(lcc_static_list))
```



# clean data

## functions

```{r}
clean_rbind = function(df){
  df %>%
    select(-.geo) %>%
    select(-`system:index`) %>%
    relocate(ID)
}
```


```{r}
clean_cbind = function(df){
  df %>%
    select(-contains(".y")) %>%
    select(-contains(".x")) 
}
```

```{r}
clean_cna = function(df){
  df %>% 
  rename(ID = id1) %>% 
  select(ID, AHM, CMD, CMD_sm, DD_0, DD_18, DD5, DD5_sm, EMT, Eref, Eref_sm, EXT, MAP, PAS, Rad_sm, SHM, Tmax_sm, Tmin_sm)
}
```
## dob

```{r}
( fisl_dob_clean = clean_rbind(fisl_dob) )
( lc_dob_clean = clean_rbind(lc_dob) )
( lcc_dob_clean = clean_rbind(lcc_dob) )
```
## indices

```{r}
( fisl_indices_clean = clean_rbind(fisl_indices) )
( lc_indices_clean = clean_rbind(lc_indices) )
( lcc_indices_clean = clean_rbind(lcc_indices) )
```
## tct

```{r}
( fisl_tct_clean = clean_rbind(fisl_tct) )
( lc_tct_clean = clean_rbind(lc_tct) )
( lcc_tct_clean = clean_rbind(lcc_tct) )
```

## treecover

```{r}
( fisl_treecover_clean = clean_rbind(fisl_treecover) )
( lc_treecover_clean = clean_rbind(lc_treecover) )
( lcc_treecover_clean = clean_rbind(lcc_treecover) )
```

## static

```{r}
( fisl_static_clean = clean_cbind(fisl_static) )
( lc_static_clean = clean_cbind(lc_static) )
( lcc_static_clean = clean_cbind(lcc_static) )
```

## CNA

```{r}
fisl_cna_clean = clean_cna(fisl_cna) 
lc_cna_clean = clean_cna(lc_cna) 
lcc_cna_clean = clean_cna(lcc_cna) 
```

# Original point files as df

```{r}
( fisl_pts_df = fisl_pts %>% st_drop_geometry() )
( lc_pts_df = lc_pts %>% st_drop_geometry() )
( lcc_pts_df = lcc_pts %>% st_drop_geometry() )
```
# Combine orginal, static, indices, tct, and tree cover by project

## Function

```{r}
combine_project_data  = function(pts, static, indices, tct, treecover, cna) {
  join_data = full_join(pts, static, by="ID") %>%
    full_join(., indices, by="ID") %>%
    full_join(., tct, by="ID") %>%
    full_join(., treecover, by="ID") %>%
    full_join(., cna, by="ID")
}
```

```{r}
( fisl_df = combine_project_data(fisl_pts_df, fisl_static_clean, fisl_indices_clean, fisl_tct_clean, fisl_treecover_clean, fisl_cna_clean) )

( lc_df = combine_project_data(lc_pts_df, lc_static_clean, lc_indices_clean, lc_tct_clean, lc_treecover_clean, lc_cna_clean) )

( lcc_df = combine_project_data(lcc_pts_df, lcc_static_clean, lcc_indices_clean, lcc_tct_clean, lcc_treecover_clean, lcc_cna_clean) )
```



# Missing data

## Project

### Disnct ids

```{r}
n_distinct(fisl_df$ID)
n_distinct(lc_df$ID)
n_distinct(lcc_df$ID)
```

### show rows with na

```{r}
( fisl_with_na <- fisl_df %>% filter(if_any(everything(), is.na)) )

( lc_with_na <- lc_df %>% filter(if_any(everything(), is.na)) )

( lcc_with_na <- lcc_df %>% filter(if_any(everything(), is.na)) )
```
### drop na



## Dob
## Function

```{r}
combine_org_dob_data  = function(pts, dob) {
  join_data = full_join(pts, dob, by="ID") 
}
```

```{r}
( fisl_dob_df = combine_org_dob_data(fisl_pts, fisl_dob_clean) )

( lc_dob_df = combine_org_dob_data(lc_pts, lc_dob_clean) )

( lcc_dob_df = combine_org_dob_data(lcc_pts_df, lcc_dob_clean) )
```


```{r}
( fisl_dob_with_na <- fisl_dob_df %>% 
    filter(if_any(everything(), is.na)) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lc_dob_with_na <- lc_dob_df %>% 
    filter(if_any(everything(), is.na))%>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lcc_dob_with_na <- lcc_dob_df %>% filter(if_any(everything(), is.na)) )
```

```{r}
fisl_dob_with_na
lc_dob_with_na
```

```{r}
sort(unique(fisl_dob_with_na$fireYr))
```

## Save to geopackage

```{r eval=FALSE, include=FALSE}
st_write(fisl_dob_with_na, "../outputs/dob_na/fisl_dob_na_2025-06-06.gpkg")
```

# DOB for FWI
