---
title: "Combustion Model"
author: "Anna Talucci"
date: "2025-04-27"
output: html_document
---


# Overview

Trouble shooting
https://stackoverflow.com/questions/30097730/error-when-using-predict-on-a-randomforest-object-trained-with-carets-train

# Library

```{r}
library(tidyverse)
library(ranger)
library(e1071)
library(arrow)
```
# Models
```{r}
above_model <- readRDS("../models/aboveground.rds")
below_model <- readRDS("../models/belowgrounds.rds")
depth_model <- readRDS("../models/burn_depth.rds")
```


# Landscapes (20) Data

```{r}
data = read_csv("../outputs/LCC_DataCombustionModel_2025-04-30.csv")
```

```{r}
head(data)
```

# Read in predictive data from stefano

```{r}
abovefed <- read_parquet("../data/TrainingDataFrame/abovefed_for_predict.parquet")
head(abovefed)
```

# Training Data

Original model training data from Stefano Potter. 

```{r}
above_train =  read.csv("../data/TrainingDataFrame/aboveground.csv", header=TRUE, sep = ",", strip.white = TRUE)
```

```{r}
below_train =  read.csv("../data/TrainingDataFrame/belowground.csv", header=TRUE, sep = ",", strip.white = TRUE)
```

```{r}
above_importance =  read.csv("../data/TrainingDataFrame/aboveground_importance.csv", header=TRUE, sep = ",", strip.white = TRUE)
```

```{r}
below_importance =  read.csv("../data/TrainingDataFrame/belowground_importance.csv", header=TRUE, sep = ",", strip.white = TRUE)
```

```{r}
above_importance
below_importance
```

# Check with AboveFed Training data

```{r}
setdiff(names(abovefed), names(data))
```

```{r}
setdiff(names(below_train), names(data))
```

```{r}
setdiff(names(above_train), names(data))
```

```{r}
setdiff(names(above_train), names(below_train))
```
```{r}
setdiff(below_importance$Variables, names(data))
```

```{r}
setdiff(above_importance$Variables, names(data))
```

```{r}
below_importance
```

# Select only importance variables and add column for combustion for training

Based on Stefano Potter's original model.

```{r}
above_train2 <- above_train %>% select(above_importance$Variables)
below_train2 <- below_train %>% select(below_importance$Variables)
above_train2$above.carbon.combusted = above_train$above.carbon.combusted
below_train2$below.ground.carbon.combusted = below_train$below.ground.carbon.combusted
```


# Training models 

Training each model with optimized parameters from Stefano Potter's ABOVEFed Model

```{r}
model_ranger_above <- ranger(above.carbon.combusted ~., data = above_train2, mtry = above_model$bestTune$mtry, splitrule = above_model$bestTune$splitrule, min.node.size = above_model$bestTune$min.node.size)
```

```{r}
model_ranger_below <- ranger(below.ground.carbon.combusted ~., data = below_train2, mtry = below_model$bestTune$mtry, splitrule = below_model$bestTune$splitrule, min.node.size = below_model$bestTune$min.node.size)
```

```{r}
rownames(below_model$importance)
below_model$bestTune$splitrule
below_model$bestTune$mtry
below_model$bestTune$min.node.size
```



# Predict combustion with all independent variables

These predictive models use all independent variables.

## Drop NA

Must remove NAs or it will not predict. 

```{r}
( df = data %>% drop_na() )
```
## select variables of interest

Following importance variables from Stefano's original model. 

```{r}
( Above = df %>% select(above_importance$Variables) )
( Below = df %>% select(below_importance$Variables) )
```

```{r}
names(Above)
names(Below)
```
## Aboveground preditions 

```{r}
pred_above = predict(model_ranger_above, Above)
```

```{r}
AbovepredResults <- data.frame(predict(model_ranger_above, Above))
AbovepredResults
```

```{r}
above_pred <- cbind(df, AbovepredResults)
above_pred
```

```{r eval=FALSE, include=FALSE}
write_csv(above_pred, "../outputs/Predict/AbovegroundCombustion_2025-04-30.csv")
```


## Below Combustion
 
```{r}
pred_below = predict(model_ranger_below, Below)
```

```{r}
BelowpredResults <- data.frame(predict(model_ranger_below, Below))

```

```{r}
below_pred <- cbind(df, BelowpredResults)
```

```{r}
below_pred
```
```{r eval=FALSE, include=FALSE}
write_csv(below_pred, "../outputs/Predict/BelowgroundCombustion_2025-04-30.csv")
```


