---
title: "LCC LC Data Explore"
author: "Anna Talucci"
date: "2025-03-19"
output: html_document
---


```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Load the LC and LCC data. Preform a brief QAQC. Generate files for acquiring spatial data sets of interest.


LCC Project Data

Field data here 

Create geopackage and shapefile for collecting spatial data

Spatial Data of interest
Fire Weather*
Long Term Climate Mean* - WorldClime version 2 or Climate NA

Stand Age https://daac.ornl.gov/cgi-bin/dsviewer.pl?ds_id=2012
Prefire NDVI
Compare Burn Depth



# Packages
```{r}
library(tidyverse)  # Data manipulation & visualization
library(data.table)  # Fast data processing
library(janitor)     # Cleaning column names
#library(summarytools) # Summary statistics
library(assertr)     # Validation checks
library(sf)
library(skimr)
library(naniar)
library(assertr)
library(naniar)
```

# Data

```{r}
LCC_Site_Data = read_csv("../data/LCC/LCC_Site_Data_01292025.csv")
LC_Plot_Final = read_csv("../data/LCC/PLOT_03262025_final.csv") # old data LC_Plot_Final_01292025.csv
```


# QAQC

## 1. View data

```{r}
head(LCC_Site_Data)
head(LC_Plot_Final)
```

```{r}
sort(unique(LCC_Site_Data$Sampling_Date))
```

### Fire names and years

```{r}
sort(unique(LC_Plot_Final$official_fire_scar))
```
```{r}
sort(unique(LCC_Site_Data$Year_of_Burn))
``` 

```{r}
sort(unique(LC_Plot_Final$recent_burn))
```
```{r}
LC_Plot_Final %>% 
  group_by(official_fire_scar, recent_burn) %>%
  summarize(n=n()) %>%
  arrange(recent_burn) 

```

```{r}
( LC_Plot_Final %>% 
  group_by(official_fire_scar, recent_burn) %>%
  summarize(n=n()) %>%
  arrange(official_fire_scar) 
  )
```


#### Fire year fix
**Note:**
Shovel creek field points should only be associated with Shovel creek fire year 2019 not 2015; need to fix typo

```{r}
LC_Plot_Final$recent_burn[LC_Plot_Final$official_fire_scar == "Shovel Creek"] <- 2019
```

```{r}
( LC_Plot_Final %>% 
  group_by(official_fire_scar, recent_burn) %>%
  summarize(n=n()) %>%
  arrange(official_fire_scar) 
  )
```
```{r}
( lc_fire_year = LC_Plot_Final %>% 
  group_by(official_fire_scar, recent_burn) %>%
  summarize(n=n()) %>%
  arrange(recent_burn) 
  )
```

```{r eval=FALSE, include=FALSE}
write.csv(lc_fire_year, "../outputs/LC_Summary_FireScar_year_2025-04-15.csv", row.names = FALSE)
```

## View data again

```{r}
glimpse(LCC_Site_Data)
glimpse(LC_Plot_Final)
```


```{r}
sort(unique(LCC_Site_Data$Researcher))
```

```{r}
LCC_Site_Data %>% group_by(Researcher)
```

## 2. Check latitude and longitude

### LCC

```{r}
LCC_Site_Data %>% filter(Longitude>0)
LCC_Site_Data %>% filter(Longitude<0)
LCC_Site_Data %>% filter(is.na(Longitude))
LCC_Site_Data %>% filter(is.na(Latitude))
```

### LC

```{r}
LC_Plot_Final %>% filter(longitude>0)
LC_Plot_Final %>% filter(longitude<0)
LC_Plot_Final %>% filter(is.na(longitude))
LC_Plot_Final %>% filter(is.na(latitude))
```

```{r}
unique(LCC_Site_Data$StudyArea) 
```

## 3. Check for duplicate rows

```{r}
sum(duplicated(LC_Plot_Final))  # Count duplicates
#df <- df[!duplicated(df), ]  # Remove duplicates
```

```{r}
sum(duplicated(LCC_Site_Data))  # Count duplicates
#df <- df[!duplicated(df), ]  # Remove duplicates
```

## 4. Check Missing values

```{r}
colSums(is.na(LCC_Site_Data))  # Count missing values per column
```

```{r}
colSums(is.na(LC_Plot_Final))  # Count missing values per column
```

```{r}
gg_miss_var(LCC_Site_Data)  # Missing values by column
gg_miss_var(LC_Plot_Final) 
```


## 5. Clean

### Clean column names 

```{r}
LCC_Site_Data <- janitor::clean_names(LCC_Site_Data)
LC_Plot_Final <- janitor::clean_names(LC_Plot_Final)
```

```{r}
LCC_Site_Data
```

### Fix longitude issue, column names, and add row id

```{r}
# Convert all positive longitudes to negative
(LCC_Site_Data <- LCC_Site_Data %>%
  mutate(longitude = ifelse(longitude > 0, -longitude, longitude)) %>%
   mutate(longitude = ifelse(study_area == "Sweden", abs(longitude), longitude)) %>%
  rename(burnYr2 = x7, burnYr3 = x8) %>% #rename columns
  mutate(row_id = row_number()) ) # add row id
```



```{r}
( LC_Plot_Final <- LC_Plot_Final %>%
  mutate(longitude = ifelse(longitude > 0, -longitude, longitude)) %>% 
  select(-site_notes) %>% 
  mutate(row_id = row_number()) )
```
```{r}
sort(unique(LC_Plot_Final$official_fire_scar))
```

```{r}
print(LCC_Site_Data)
print(LC_Plot_Final)
```


## 6. Data Checks

### Check categorical variables

```{r eval=FALSE, include=FALSE}
# Function to summarize categorical variables
summarize_categorical <- function(df) {
  df %>%
    select(where(~ is.character(.) | is.factor(.))) %>%
    map(~ list(
      unique_values = unique(.),
      count = table(.),
      missing = sum(is.na(.))
    ))
}

# Apply function
summarize_categorical(LC_Plot_Final)
summarize_categorical(LCC_Site_Data)
# Print summary for each categorical variable


```

### Check Numeric Data 

```{r}
skim(LC_Plot_Final) %>%
   mutate(across(where(is.numeric), ~ round(., 2)))
```

```{r}
skim(LCC_Site_Data) %>%
   mutate(across(where(is.numeric), ~ round(., 2)))
```



```{r}
head(LC_Plot_Final)
head(LCC_Site_Data)
```

```{r}
summary(LC_Plot_Final)
summary(LCC_Site_Data)
```

# Write

```{r}
write.csv(LC_Plot_Final, "../outputs/LCC_cleaned/csv/LC_Plot_cleaned_2025-04-16.csv")
write.csv(LCC_Site_Data, "../outputs/LCC_cleaned/csv/LCC_Site_cleaned_2025-04-16.csv")
```