---
title: "Figures FWI Variables"
author: "Anna Talucci"
date: "2025-08-11"
output: html_document
---


# Figures 

## Plot themes

```{r}
comp_theme = theme_bw() + theme(legend.position = "bottom") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"))
```

## Color coding
'#256676', Baltzer
'#7cd3eb', Bergeron -
'#3a427d', Diaz Veraverbeke 
'#956bed', Dominque Arsonault 
'#ffa8ff', Gundal/Nisson 
'#8a2f6b', Hayes
'#f75ef0', Hung & Natali 
'#1932bf', Kristen Reid
'#6e9df7', Lucas Brehaut 
'#45deb2', Metcalfe 
'#069668', Pare 
'#a5e841', Parisien -
'#0b5313', Saperstein
'#cadba5' Walker

## fwi

```{r}
( fwi_plot = ggplot(fwi, aes(x=PI, y=fire_weather_index, color = PI)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Fire Weather Index") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-29_fwi.png", plot = fwi_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```


## keetch_byram_drought_index
keetch_byram_drought_index 0 to 800. A value of 0 indicates no moisture deficit, while 800 represents the maximum possible drought. 
```{r}
( drought_plot = ggplot(fwi, aes(x=PI, y=keetch_byram_drought_index, color = PI)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Keetch Byram drought index") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-29_drought.png", plot = drought_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```




## fine_fuel_moisture_code

```{r}
( ffm_plot = ggplot(fwi, aes(x=PI, y=fine_fuel_moisture_code, color = PI)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Fine fuel moisture code") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-29_FFMC.png", plot = ffm_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

## initial_fire_spread_index

```{r}
( isi_plot = ggplot(fwi, aes(x=PI, y=initial_fire_spread_index, color = PI)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Initial spread index") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-29_ISI.png", plot = isi_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```
