---
title: "Modis Day of Burn"
author: "Anna Talucci"
date: "2025-04-17"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Take Day of Burn day of year day and convert to YYYY-MM-DD for EE script for extracting FWI.

Input data is from Earth Engine output

**PART OF WORK FLOW 8/11/2025**
# packages

```{r}
library(tidyverse)
library(sf)
```

# Data

## point data

```{r}
lc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lc_pts_WGS84.shp", "2025-08-11_lc_pts_WGS84")
```

```{r}
lcc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lcc_pts_WGS84.shp", "2025-08-11_lcc_pts_WGS84")
```

## Day of bun

```{r}
lc_files <- list.files("../data/2025-08-08_SpatialData/DoBModis/lc", pattern='.csv$', full.names=TRUE)
```

```{r}
lcc_files <- list.files("../data/2025-08-08_SpatialData/DoBModis/lcc", pattern='.csv$', full.names=TRUE)
```

```{r}
lc_files
```

```{r}
lcc_files
```

# Read into dataframe

```{r}
lc_df <- map_dfr(lc_files, readr::read_csv, show_col_types = FALSE)
```



```{r}
lcc_df <- map_dfr(lcc_files, readr::read_csv, show_col_types = FALSE)
```

```{r}
lcc_df
```

```{r}
( lc_dob = lc_df %>% mutate(date= as.Date(BurnDate-1, origin=paste0(fireYr, "-01-01")), 
              month= strftime(date, "%m"), 
              day=strftime(date,"%d"))  %>%
  rename(dob=BurnDate) %>%
  relocate(ID, date) %>%
  dplyr::select(ID, dob, date)
  
)
```
```{r}
lcc_df
```

```{r}
( lcc_dob = lcc_df %>% mutate(date= as.Date(BurnDate-1, origin=paste0(fireYr, "-01-01")), 
              month= strftime(date, "%m"), 
              day=strftime(date,"%d"))  %>%
  rename(dob=BurnDate) %>%
  relocate(ID, date) %>%
  dplyr::select(ID, dob, date)
  
)
```

# Join to points

```{r}
lc_pts
lcc_pts
```

```{r}
( lc_dob_pts = lc_pts %>% 
    full_join(lc_dob, by='ID') %>%
    mutate(date = as.character(date)))
```

```{r}
( lcc_dob_pts = lcc_pts %>% 
    full_join(., lcc_dob, by="ID") %>% 
    mutate(date = as.character(date)))
```

```{r}
glimpse(lc_dob_pts)
```

# Look at dob NA

```{r}
missing_dob = lc_dob_pts %>% filter(is.na(dob))
```

```{r}
missing_dob
unique(missing_dob$PI)
sort(unique(missing_dob$fireYr))
```
```{r}
lcc_missing_dob = lcc_dob_pts %>% filter(is.na(dob))
```

```{r}
lcc_missing_dob
unique(lcc_missing_dob$PI)
sort(unique(lcc_missing_dob$fireYr))
```

```{r}
lcc_missing_dob %>% filter(PI=="Parisien")
```

```{r}
(lc_drop_na = lc_dob_pts %>% filter(!is.na(dob)))
```

```{r}
(lcc_drop_na =lcc_dob_pts %>% filter(!is.na(dob)))
```

# look at fire years

```{r}
sort(unique(lc_dob_pts$fireYr))
```

```{r}
sort(unique(lcc_dob_pts$fireYr))
```

```{r}
lcc_drop_na %>% group_by(fireYr) %>% summarise()
```
```{r}
lcc_dob_pts
lc_dob_pts
```

# Print calendar dates for EE

Get calendar date with quotes and separated by comma to add to GEE code. Copy printed list into EE code `BurnDateFWIExtract`

```{r}
cat(paste0(sprintf("'%s'", sort(unique(lcc_dob_pts$date))), collapse = ", "))
```

```{r}
cat(paste0(sprintf("'%s'", sort(unique(lc_dob_pts$date))), collapse = ", "))
```

# save all points

```{r eval=FALSE, include=FALSE}
st_write(lc_dob_pts, "../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lc_pts_dob.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(lcc_dob_pts, "../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lcc_pts_dob.shp", driver="ESRI Shapefile")
```

# save shapefile for EE

```{r eval=FALSE, include=FALSE}
st_write(lc_drop_na, "../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lc_pts_dob_dropNa.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(lcc_drop_na, "../outputs/LCC_cleaned/points/shapefiles/2025-08-11_lcc_pts_dob_dropNa.shp", driver="ESRI Shapefile")
```
