---
title: "Organize Combustion Model Data"
author: "Anna Talucci"
date: "2025-04-27"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```


# Overview
 LCC & LC clean and combine data for combustion model.
 
 Create one data frame with all sampled points and variables. The data frame will be subset in the model script.

# Packages

```{r}
library(tidyverse)
```

# Data

```{r}
pts = st_read("../outputs/LCC_cleaned/points/shapefiles/lc_lcc_2025-04-24.shp", "lc_lcc_2025-04-24")
```
```{r}
( sites = pts %>% st_drop_geometry() )
```
## FWI

```{r}
gfwed =read_csv("../outputs/LCC_cleaned/csv/gfwed_2025-04-30.csv")
```
```{r}
fwi = read_csv("../outputs/LCC_cleaned/csv/fwi_2025-04-29.csv")
```

```{r}
( fwi = fwi %>% 
    rename(BUI = build_up_index, DC= drought_code, DMC = duff_moisture_code, DOB_lst = dob, FFMC = fine_fuel_moisture_code, FWI = fire_weather_index, ISI = initial_fire_spread_index) %>%
    dplyr::select(row_id, siteplot, date, DOB_lst, BUI, DC, DMC, FFMC, FWI, ISI ))
```
 [1] "BUI"               "DC"                "DMC"               "DOB_lst"          
 [5] "DSR"               "Date"              "FFMC"              "FWI"              
 [9] "ID1"               "ID2"               "ISI"               "Relative.humidity"
[13] "Temperature"       "VPD"               "Wind.speed"  

```{r}
gfwed = gfwed %>% rename(Relative.humidity = rh, Temperature=t, VPD=vpd, Wind.speed = wspeed)
```


**NOTE Current FWI DATA is missing RH, Temp, VPD, Wind speed, and DSR.**

```{r}
fwi 
```

```{r}
min(fwi$FWI)
max(fwi$FWI)
```

###Join

```{r}
( fwi1 = full_join(fwi, gfwed, by = c("row_id", "siteplot"))
)
```


## Landsat
### indices
```{r}
ls_indices =  list.files(path="../data/landsatIndices",pattern='*.csv', full.names = TRUE) %>% 
  purrr::map(., read_csv) %>% 
  map_dfr(., ~ .x ) %>%
  bind_rows() %>%
  dplyr::select(-`system:index`, -`.geo`) %>%
  rename(NDVI=dNDVI, NDII=dNDII)
```
```{r}
ls_indices
```

### TCT

```{r}
tct =  list.files(path="../data/TCT",pattern='*.csv', full.names = TRUE) %>% 
  purrr::map(., read_csv) %>% 
  map_dfr(., ~ .x ) %>%
  bind_rows() %>%
  dplyr::select(row_id, siteplot, wetness, greenness, brightness) 
```

```{r}
tct
```

### Tree Cover

```{r}
treecover =  list.files(path="../data/treeCover",pattern='*.csv', full.names = TRUE) %>% 
  purrr::map(., read_csv) %>% 
  map_dfr(., ~ .x ) %>%
  bind_rows() %>%
  dplyr::select(row_id, siteplot, tree_canopy_cover) %>%
  rename(Tree.cover=tree_canopy_cover) 
```

```{r}
treecover 
```

###Join

```{r}
( landsat = full_join(ls_indices, tct, by = c("row_id", "siteplot")) %>%
    full_join(., treecover, by = c("row_id", "siteplot")))
```

```{r}
sort(unique(landsat$PI))
```
## Static

```{r}
terrain = read_csv("../data/static/staticTerrain.csv")
soil = read_csv("../data/static/staticSoil250_LCC.csv")
soilPfi = read_csv("../data/static/staticSoilPFI_LCC.csv")
sveg = read_csv("../data/static/static1Vegetation_LCC.csv")
```

```{r}
terrain = terrain %>%
  dplyr::select(-`system:index`, -`.geo`) 
soil = soil %>%
  dplyr::select(row_id, siteplot, BD_30, Clay_30, SOC_30, Sand_30, Silt_30, pH_30)
soilPfi = soilPfi %>%
  dplyr::select(row_id, siteplot, PFI)
sveg = sveg %>%
  dplyr::select(row_id, siteplot, BS, DEC, GRSH, JP, NV, OCON, PI, WS)
```



### Join

```{r}
( static = full_join(terrain, sveg, by = c("row_id", "siteplot")) %>%
    full_join(., soil, by = c("row_id", "siteplot")) %>%
    full_join(., soilPfi, by = c("row_id", "siteplot"))  %>%
    select(-contains(".y")) %>%
    rename_at(vars(contains(".x")), ~str_remove(., ".x"))
  )
```

```{r}
sort(unique(static$PI))
```




## Climate NA

```{r}
lc_1971 = read_csv("../data/climateNA/climateNA_LC_2025-04-17_Normal_1971_2000MSY.csv")
lc_1991 = read_csv("../data/climateNA/climateNA_LC_2025-04-17_Normal_1991_2020MSY.csv")
lcc_1971 = read_csv("../data/climateNA/climateNA_LCC_2025-04-17_Normal_1971_2000MSY.csv")
lcc_1991 = read_csv("../data/climateNA/climateNA_LCC_2025-04-17_Normal_1991_2020MSY.csv")
```
  
```{r}
( climateNa_1971 = bind_rows(lc_1971, lcc_1971) %>% 
  dplyr::select(id1, id2, AHM, CMD, CMD_sm, DD_0, DD_18, DD5, DD5_sm, EMT, Eref, Eref_sm, EXT, MAP, PAS, Rad_sm, SHM, Tmax_sm, Tmin_sm) %>%
  rename(ID1=id1, ID2=id2, CNA_AHM=AHM, CNA_CMD=CMD, CNA_CMD_5_8 = CMD_sm, CNA_DD_0=DD_0, CNA_DD_18=DD_18, CNA_DD5=DD5, CNA_DD5_5_8=DD5_sm, CNA_EMT=EMT, CNA_Eref=Eref, CNA_Eref_5_8=Eref_sm, CNA_EXT=EXT, CNA_MAP=MAP, CNA_PAS=PAS, CNA_Rad_5_8=Rad_sm, CNA_SHM=SHM, CNA_Tmax_5_8=Tmax_sm, CNA_Tmin_5_8=Tmin_sm) %>%
    rename(row_id = ID1)
)
```

```{r}
( cna = sites %>% full_join(climateNa_1971, by="row_id")
)
```

```{r}
sort(unique(cna$PI))
```

```{r}
cna %>% filter(if_any(everything(), is.na)) 
```

# Check number of IDs

```{r}
length(unique(landsat$row_id))
length(unique(static$row_id))
length(unique(fwi$row_id))
length(unique(climateNa_1971$row_id))
```





# Combine to Datasets

```{r}
( allData = full_join(static, landsat, by = c("row_id", "siteplot")) %>%
  full_join(., fwi1, by = c("row_id", "siteplot")) %>%
   full_join(., cna, by = c("row_id", "siteplot")) %>%
    select(-contains(".y")) %>%
    select(-c(45, 52:56)) %>%
    
    rename(PI=PI.x, area=area.x, fireYr=fireYr.x, lat=lat.x, lon=lon.x, proj=proj.x, date=date.x)
)
```


```{r}
allData %>% filter(PI=="Saperstein")
```

# Fill in stand age

```{r}
( allData1 = allData %>%
mutate(Tree.cover = case_when(
    is.na(Tree.cover) & PI == "Diaz-Veraverbeke" ~ 0,
    TRUE ~ Tree.cover  # keep existing value otherwise
  )) %>% 
  mutate(Tree.cover = case_when(
    is.na(Tree.cover) & PI == "Hung and Natali" ~ 0,
    TRUE ~ Tree.cover  # keep existing value otherwise
  ))
)

```

## Check NAs
There are several NAs for the FWI Data

```{r}
allData[rowSums(is.na(allData)) > 0,]
```

```{r}
( datawithNA = allData %>% filter(if_any(everything(), is.na)) )
```
```{r}
write_csv(datawithNA, "../outputs/LCC_DataCombustionModelWithNA_2025-04-30.csv")
```
```{r}
( naFree = allData %>% 
    drop_na() )
```

```{r}

naFree %>% group_by(fireYr, PI) %>% summarise(n = n())

```


# Write to csv

```{r}
write_csv(allData1, "../outputs/LCC_DataCombustionModel_2025-04-30.csv")
```


**THE END**