---
title: "Organize Single Dataframe"
author: "Anna Talucci"
date: "2025-04-22"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Combine spatial data in to single dataframe


color help
http://vrl.cs.brown.edu/color
# Packages

```{r}
library(tidyverse)
```

# Plot themes

```{r}
comp_theme = theme_bw() + theme(legend.position = "bottom") +  
  theme(panel.grid.major = element_blank()) +
    theme(axis.title.y = element_text(size = 11, hjust = 0.5, vjust = 1.1),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.text.y = element_text(size = 10, color = "black"))
```

# Data

## All site data

```{r}
allData = read_csv("../data/LCC/LCC_Site_clean.csv")
```

## Spatial Data

### Forest Age

```{r}
lcc_forest_age = read_csv("../outputs/LCC_cleaned/csv/LCC_stand_age_2025-04-22.csv")
lc_forest_age = read_csv( "../outputs/LCC_cleaned/csv/LC_stand_age_2025-04-22.csv")
```

### Climate

```{r}
lc_climate = read_csv("../outputs/extractSpatialData/lc_worldclim2_monthlyClimate_2025-04-17.csv")  
lcc_climate = read_csv("../outputs/extractSpatialData/lcc_worldclim2_monthlyClimate_2025-04-17.csv")   
```

# View

```{r}
allData
```

```{r}
lc_forest_age
lcc_forest_age
```

```{r}
lc_climate
lcc_climate
```



# Split all data into LC and LCC

```{r}
( lc = allData %>% filter(project=="LC") )
( lcc = allData %>% filter(project=="LCC") )

```

# Reduce columns on Spatial data

```{r}
lc_age = lc_forest_age %>% dplyr::select(fire:unique_site_id, row_id, stand_age)
lcc_age = lcc_forest_age %>% dplyr::select(researcher, site, plot, forest_age) %>% rename(stand_age = forest_age)
```

```{r}
( lc_wc2 = lc_climate %>% dplyr::select(fire:unique_site_id,  wc2.1_30s_tmin_01:wc2.1_30s_bio_9) )
( lcc_wc2 = lcc_climate %>% dplyr::select(researcher, site, plot, wc2.1_30s_tmin_01:wc2.1_30s_bio_9)  )
```


```{r}
unique(lcc_wc2$siteplot)
unique(lcc$siteplot)
```

# Combine


```{r eval=FALSE, include=FALSE}
( lcc_join = lcc %>% 
  full_join(., lcc_wc2, by = "siteplot") %>%
  full_join(., lcc_age, by = "siteplot") %>%
    drop_na(siteplot))
```

```{r eval=FALSE, include=FALSE}
full_join(lcc %>% group_by(site) %>% mutate(id = row_number()),
          lcc_wc2 %>% group_by(site) %>% mutate(id = row_number()),
          by = c("id", "site"))
```


# Data for plots
researcher, study_area/fire, site/plot
```{r}
( lc_age = lc_forest_age %>% 
  mutate(researcher="Walker") %>% 
  rename(study_area = fire, forest_age = stand_age) %>% 
  unite("siteplot", site:plot, remove = FALSE) %>%
  dplyr::select(researcher, study_area, siteplot, forest_age) )
```

```{r}
( lcc_age = lcc_forest_age %>% 
  rename(siteplot = site) %>%
  dplyr::select(researcher, study_area, siteplot, forest_age) )
  
```

```{r}
( forest_age = bind_rows(lc_age, lcc_age) %>% drop_na(researcher))
```

```{r}
( climate_LC = lc_climate %>% 
  mutate(researcher="Walker") %>% 
  rename(study_area = fire) %>% 
  unite("siteplot", site:plot, remove = FALSE) %>%
  dplyr::select(researcher, study_area, siteplot, wc2.1_30s_tmin_01:wc2.1_30s_bio_9) )
```

```{r}
( climate_LCC = lcc_climate %>% 
  rename(siteplot = site) %>%
  dplyr::select(researcher, study_area, siteplot, wc2.1_30s_tmin_01:wc2.1_30s_bio_9) )
```

```{r}
( climate = bind_rows(climate_LC, climate_LCC) %>% drop_na(researcher))
```


# Box Plots

## Color coding
'#256676', Baltzer
'#7cd3eb', Bergeron
'#3a427d', Diaz Veraverbeke
'#956bed', Dominque Arsonault
'#ffa8ff', Gundal/Nisson
'#8a2f6b', Hayes
'#f75ef0', Hung & Natali
'#1932bf', Kristen Reid
'#6e9df7', Lucas Brehaut
'#45deb2', Metcalfe
'#069668', Pare
'#a5e841', Parisien
'#0b5313', Saperstein
'#cadba5' Walker

## Forest Age

```{r}
( forest_age_plot = ggplot(forest_age, aes(x=researcher, y=forest_age, color = researcher)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#7cd3eb', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#a5e841', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Stand Age (satillite derived)") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-24_forestAge.png", plot = forest_age_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

## Climate

```{r}
( precip_wet_plot = ggplot(climate, aes(x=researcher, y=wc2.1_30s_bio_13, color = researcher)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#7cd3eb', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#a5e841', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Precipitation wettest month (WorldClim2)") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-24_precipWettest.png", plot = precip_wet_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( precip_dry_plot = ggplot(climate, aes(x=researcher, y=wc2.1_30s_bio_14, color = researcher)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#7cd3eb', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#a5e841', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Precipitation driest month (WorldClim2)") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-24_precipDriest.png", plot = precip_dry_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( temp_warm_plot = ggplot(climate, aes(x=researcher, y=wc2.1_30s_bio_10, color = researcher)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#7cd3eb', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#a5e841', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Mean temperature warmest quarter (WorldClim2)") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-24_tempWarmestquarter.png", plot = temp_warm_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```

```{r}
( temp_Max_plot = ggplot(climate, aes(x=researcher, y=wc2.1_30s_bio_5, color = researcher)) +
  geom_boxplot( linewidth = .3 ) + #,outlier.shape = NA, size = 3, shape = 21
  geom_jitter( size = 1, shape = 21) +
      scale_color_manual(values=c('#256676', '#7cd3eb', '#3a427d', '#956bed', '#ffa8ff', '#8a2f6b', '#f75ef0', '#1932bf', '#6e9df7', '#45deb2', '#069668', '#a5e841', '#0b5313', '#cadba5'), name="") +
    stat_summary(fun=mean, geom="point", shape=23, size=2, color="tomato") +

  labs(x="Researcher", y="Max temperature of warmest month (WorldClim2)") +
  coord_flip() +

  comp_theme +
    theme(legend.position = "none",
          axis.text.y = element_text(size=8),
          axis.text.x = element_text(size=8),
          strip.background = element_rect( color="white", fill="white", size=1.5, linetype="solid" )) +
    theme(strip.text.x = element_text(size = 9))
)
```

### Save Figures
```{r eval=FALSE, include=FALSE}
ggsave("../figures/boxplots/2025-04-24_tempMaxMonth.png", plot = temp_Max_plot,  width = 6, height =5, units = c("in"), dpi=600, bg = "white" )
```