---
title: "Spatialize Data"
author: "Anna Talucci"
date: "2025-04-14"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Use cleaned qaqc data and spatialize



# Packages
```{r}
library(tidyverse)  # Data manipulation & visualization

library(sf)
library(mapview)
```


# Data

```{r}
lc = read_csv("../outputs/LCC_cleaned/csv/LC_Plot_cleaned_2025-04-16.csv")
lcc = read_csv("../outputs/LCC_cleaned/csv/LCC_Site_cleaned_2025-04-16.csv")
```

# Reorganize data and Spatialize

This reformats data for Climate NA and spatializes data with limited attributes to use for extracting various spatial data products. 



# Spatialize Data

```{r}
lc
lcc
```
### Function to spatialize

```{r}
df_to_sf_wgs84 <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4326, remove = FALSE) # X. Walker states crs epsg:3338 but it ends up in the ocean 
}
```

```{r}
df_to_sf_NAD83 <- function(x){
  st_as_sf(x, coords = c("longitude","latitude"), crs = 4269, remove = FALSE) # X. Walker states crs epsg:3338 but it ends up in the ocean 
}
```

### points

```{r}
( lc_pts_wgs84 = lc %>% 
  #select(row_id, fire, site, plot, unique_site_id, latitude, longitude, recent_burn) %>% 
    df_to_sf_wgs84() 
)
```

```{r}
( lc_pts_NAD83 = lc %>% 
  #select(row_id, fire, site, plot, unique_site_id, latitude, longitude, recent_burn) %>% 
    df_to_sf_NAD83() 
)
```

```{r}
lc_pts_NAD83_wgs84 = st_transform(lc_pts_NAD83, crs = 4326)
```

```{r}
( lcc_pts_wgs84 = lcc %>% 
    #drop_na(latitude) %>%
    df_to_sf_wgs84() 
)
```

```{r}
( lcc_pts_NAD83_wgs84 = lcc %>% 
    #drop_na(latitude) %>%
    df_to_sf_NAD83() %>%
    st_transform(., crs = 4326)
)
```

```{r}
( ee_lcc_pts_wgs84 = lcc_pts_wgs84 %>% 
    dplyr::select(row_id, researcher, study_area, site, sampling_date, year_of_burn, latitude, longitude) %>%
    rename(PI = researcher, area= study_area, dateSmp = sampling_date, burnYr = year_of_burn, lat = latitude, lon = longitude)
)
```
```{r}
ee_lcc_pts_NAD83_wgs84 = lcc_pts_NAD83_wgs84 %>% 
    dplyr::select(row_id, researcher, study_area, site, sampling_date, year_of_burn, latitude, longitude) %>%
    rename(PI = researcher, area= study_area, dateSmp = sampling_date, burnYr = year_of_burn, lat = latitude, lon = longitude)
```

```{r}
ee_lc_pts_NAD83_wgs84 = lc_pts_NAD83_wgs84 %>% 
  dplyr::select(row_id, fire, site, plot, unique_site_id, latitude, longitude, recent_burn, official_fire_scar) %>% 
  rename(siteID = unique_site_id, lat = latitude, lon = longitude, fireYr = recent_burn, fireNm = official_fire_scar)
```

# View points

```{r}
mapview(lc_pts_wgs84)
```

```{r}
mapview(lc_pts_NAD83)
```

```{r}
mapview(lc_pts_NAD83_wgs84)
```

```{r}
mapview(lcc_pts_wgs84)
```

```{r}
mapview(lcc_pts_NAD83)
```

# Save points with all attributes

## geopackage

```{r eval=FALSE, include=FALSE}
st_write(lc_pts_NAD83_wgs84, "../outputs/LCC_cleaned/points/geopackage/lc_pts_NAD83_wgs84_2025-04-16.gpkg", )
```

```{r eval=FALSE, include=FALSE}
st_write(lcc_pts_wgs84, "../outputs/LCC_cleaned/points/geopackage/lcc_pts_wgs84_2025-04-16.gpkg")
```

## shapefiles
```{r eval=FALSE, include=FALSE}
st_write(ee_lc_pts_NAD83_wgs84, "../outputs/LCC_cleaned/points/shapefiles/lc_pts_NAD83_wgs84_2025-04-16.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(ee_lcc_pts_wgs84, "../outputs/LCC_cleaned/points/shapefiles/lcc_pts_wgs84_2025-04-16.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(ee_lcc_pts_NAD83_wgs84, "../outputs/LCC_cleaned/points/shapefiles/lcc_pts_NAD83_wgs84_2025-04-16.shp", driver="ESRI Shapefile")
```

### Spatialize to points for FWI

```{r}
lc_pts_NAD83_wgs84
lcc_pts_wgs84
```


```{r}
lcc_pts_wgs84 %>% 
  group_by(study_area, year_of_burn) %>% 
  summarise(n=n())
```

```{r}
lc_pts_NAD83_wgs84 %>% 
  group_by(official_fire_scar, recent_burn) %>% 
  summarise(n=n()) %>%
  arrange(recent_burn)
```

```{r}
unique(lc_pts_NAD83_wgs84$official_fire_scar)
```

```{r}
sort(unique(lc_pts_NAD83_wgs84$recent_burn))
```
```{r}
 lc_pts_NAD83_wgs84 %>% 
  filter(official_fire_scar %in% c("Bearnose Hill", "Chitanana", "Douglas", "Frozen Calf", "U.S. Creek")) %>%
  group_by(official_fire_scar, recent_burn) %>% 
  summarise(n = n())
```

```{r}
lc_pts_NAD83_wgs84 %>% 
  distinct(official_fire_scar, recent_burn) %>% 
  group_by(recent_burn) %>% 
  summarise(fireNames = paste(unique(official_fire_scar), collapse = ", "))
```


```{r}
LC_Plot_Final_pts = LC_Plot_Final %>% 
  dplyr::select(row_id, site, plot, unique_site_id, latitude, longitude, recent_burn) %>%
  unite(siteID, c("site", "plot", "unique_site_id")) %>%
  df_to_sf() %>% 
  st_transform(., crs = 4326) %>%
  mutate(dataset = "LC_Plot_Final")  %>%
  rename(burnYr = recent_burn) %>%
  mutate(burnYr = as.character(burnYr))
```

```{r}

LCC_Site_Data_pts = LCC_Site_Data %>% 
  dplyr::select(row_id, site, latitude, longitude, year_of_burn) %>%
  unite(siteID, c("site", "row_id"), sep = "_", remove = FALSE)  %>% 
  drop_na(latitude, longitude) %>%
  df_to_sf() %>% 
  st_transform(., crs = 4326) %>%
  mutate(dataset = "LCC_Site_Data") %>%
  rename(burnYr = year_of_burn) %>%
  mutate(burnYr = as.character(burnYr))
```

```{r}
LC_Plot_Final_pts
LCC_Site_Data_pts
```



# Save

```{r eval=FALSE, include=FALSE}
st_write(LC_Plot_Final_pts, "../outputs/LCC_cleaned/points/LC_Plot_ptsWGS84_2025-04-02.shp", driver="ESRI Shapefile")
```

```{r eval=FALSE, include=FALSE}
st_write(LCC_Site_Data_pts, "../outputs/LCC_cleaned/points/LCC_Site_ptsWGS84_2025-04-02.shp", driver="ESRI Shapefile")
```

## Filter reburn data

Reburn data for FiSL biophysical properties research

```{r}
reburnData = LC_Plot_Final %>% 
  dplyr::select(row_id, fire, site, plot, unique_site_id, latitude, longitude, recent_burn, second_burn, triple_burn, reburn_age_original_estimate, reburn_class_final)
```

```{r}
write.csv(reburnData, "../outputs/LCC_cleaned/csv/LC_Plot_ReburnData_2025-04-02.csv")
```

## Climate NA data formating

```{r}
lcc %>% group_by(researcher, study_area, site) %>% summarise(n=n())
```

```{r}
climateNA_lc = lc %>% 
  dplyr::select(row_id, site, plot, unique_site_id, latitude, longitude) %>%
  unite(siteID, c("site", "plot", "unique_site_id")) %>%
  rename(id1 = row_id, id2 = siteID, lat = latitude, long = longitude) 

```

```{r}
climateNA_lcc  = lcc %>% 
  filter(study_area != "Sweden") %>%
  dplyr::select(row_id, researcher, site, latitude, longitude) %>%
  unite(siteID, c("site", "row_id"), sep = "_", remove = FALSE) %>%
  rename(id1 = row_id, id2 = siteID, lat = latitude, long = longitude) %>%
  select(id1, id2, lat, long)
```

```{r}
write.csv(climateNA_lc, "../outputs/LCC_cleaned/csv/climateNA_LC_2025-04-15.csv", row.names = FALSE)
write.csv(climateNA_lcc, "../outputs/LCC_cleaned/csv/climateNA_LCC_2025-04-15.csv", row.names = FALSE)
```

